### 2023-12-15 paper.tex

\documentclass{article}
\usepackage{arxiv}
\input{common}

\title{%
    Your Title Goes Here...}
\author{%
    Your Name Here\thanks{%
        Equal contribution.
        Correspondence to Xitong Gao
        (\email{xt.gao@siat.ac.cn}).
    }\hspace{0.4em}\authornote{1,2},
    Xitong Gao\samethanks[1]\authornote{1},
    Kejiang Ye\authornote{1},
    Cheng-Zhong Xu\authornote{3} \\
    \authornote{1}\,%
        Shenzhen Institutes of Advanced Technology,
        Chinese Academy of Sciences, China. \\
    \authornote{2}\,%
        University of Chinese Academy of Sciences, China. \\
    \authornote{3}\,%
        University of Macau, Macau S.A.R., China.
}
\date{}

\begin{document}

\maketitle

\input{abstract}
\input{intro}
\input{related}
\input{method}
\input{results}
% \FloatBarrier
\input{conclusion}
% \clearpage
\bibliographystyle{plainnat}
\bibliography{references}
\newpage
\appendix
% \onecolumn
\input{appendix}

\end{document}