### 2024-01-11 conclusion.tex
Update template

\section{Conclusion}\label{sec:conclusion}

Conclusion goes here\dots{}
This section should provide a brief summary of the paper.
It should also highlight the contributions of the paper,
and discuss the implications of it.

\section{Reproducibility Statement}\label{sec:reproducibility}

Some conferences require a reproducibility statement.
The reproducibility statement
should include the following:
\begin{itemize}
    \item A list of the computational resources used in the work.

    \item Hyperparameter settings,
    can reference relevant sections in Appendix.

    \item A link to access the code
    with a complete description of how to run it.
    The link should be anonymous for double-blind submissions.
\end{itemize}

\section{Ethics Statement}\label{sec:ethics}

Some conferences require an ethics statement.
The ethics statement should include the following:
\begin{itemize}
    \item A discussion of ethical issues raised by the study,
    if applicable.
\end{itemize}

\section{Limitations}\label{sec:limitations}

Some conferences require a limitations section.
The limitations section should include the following:
\begin{itemize}
    \item A discussion of the limitations of the study.
    \item A discussion of how the limitations
    could be addressed in future work.
\end{itemize}

\section*{Acknowledgements}\label{sec:acknowledgements}

This is a section to be added after paper acceptance.
Double-blind submissions
should not include this section.

% An example snippet of acknowledgements:
\begin{comment}
    This research is supported
    by National Natural Science Foundation of China
    (\numero{62376263}),
    and Basic Research Program of Shenzhen
    (\numero{JCYJ20230807140507015}).
\end{comment}