### 2024-11-04 paper.tex
Initial commit

\documentclass{article}
\usepackage{arxiv}
\input{common}

\title{%
    Your Title Goes Here...}
\author{%
    Your Name Here\hspace{0.4em}\authornote{1} \\
    \authornote{1}\,%
        Your affiliation here... \\
}
\date{}

\begin{document}

\maketitle

\input{abstract}
\input{intro}
\input{related}
\input{method}
\input{results}
\input{conclusion}
% \clearpage
\bibliographystyle{plainnat}
\bibliography{references}
\newpage
\appendix
% \onecolumn
\input{appendix}

\end{document}