### 2024-11-04 conclusion.tex
Update template to CVPR 2025

\section{Conclusion}\label{sec:conclusion}

\section{Reproducibility Statement}\label{sec:reproducibility}

\section{Ethics Statement}\label{sec:ethics}

\section{Limitations}\label{sec:limitations}