### 2024-11-04 paper.tex

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage[review]{cvpr}
\definecolor{cvprblue}{rgb}{0.21,0.49,0.74}
\usepackage[pagebackref,breaklinks,colorlinks,allcolors=cvprblue]{hyperref}
\def\paperID{*****} % TODO: Enter the Paper ID here
\def\confName{CVPR}
\def\confYear{2025}

\input{common}

\title{%
    Your Title Goes Here...}
\author{%
    Your Name Here\hspace{0.4em}\authornote{1} \\
    \authornote{1}\,%
        Your affiliation here... \\
}
\date{}

\begin{document}

\maketitle

\input{abstract}
\input{intro}
\input{related}
\input{method}
\input{results}
\input{conclusion}
{
    \small
    \bibliographystyle{ieeenat_fullname}
    \bibliography{references}
}
\newpage
\appendix
% \onecolumn
\input{appendix}

\end{document}